---
title: "ps_4"
author: "Henry Zhu"
date: "February 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
library(gt)
library(devtools)
library(tidyverse)
library(readr)
library(janitor)
nc_poll<- read_csv("ps_4_elections-poll-nc09-3.csv")%>% clean_names()
dem<- nc_poll%>%
  filter(response=="Dem")
rep<- nc_poll%>%
  filter(response=="Rep")
und<- nc_poll%>%
  filter(response=="Und")
gendiff<-nc_poll%>% filter(gender!= gender_combined)
whitediff <-nc_poll%>%select(race_eth, file_race_black)%>%filter(race_eth=="White",race_eth!=file_race_black)
```
There were `r nrow(dem)` respondents who supported the Democratic candidate.

There were `r (nrow(rep)-nrow(und))` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables (gender and gender_combined). There are `r nrow(gendiff)` individuals for whom these variables have different values.

There are `r nrow(whitediff)` respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.